| Base de datos de abejas ibéricas \n| Iberian bees database\n

### 2022-07-01 Manuscript/Figures/Figure_1.Rmd
manuscript update

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',out.width="115%"}
library(sf)
library(raster)
library(rasterVis)
library(tidyverse)
library(data.table)
data <- read.table("../../Data/iberian_bees.csv.gz", header = T, quote = "\"", sep = ",")

@@

# install.packages("ggplot2")
library(ggplot2)
library(viridis)
library(giscoR) #map of europe
library(ggspatial)

#gisco
#gisco_countrycode
eu2016 <- c(gisco_countrycode[gisco_countrycode$eu, ]$CNTR_CODE)

#Select data
europe_map <- gisco_get_nuts(
  year = "2016",
  epsg = "3035",
  resolution = "3",
  nuts_level = "2",
  country = eu2016) %>%
  sf::st_transform(crs = 4326)

# Borders
borders <- gisco_get_countries(
  epsg = "3035",
  year = "2016",
  resolution = "3",
  country = eu2016) %>%
  sf::st_transform(crs = 4326)

vars <- c("ES",  "PT")
#Extract counties of interest
euro_map <- filter(europe_map, CNTR_CODE %in% vars)

@@

library(ggsn)

>>>>

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',out.width="115%"}
library(sf)
library(raster)
library(rasterVis)
library(tidyverse)
library(data.table)
library(ggplot2)
library(viridis)
library(ggspatial)
library(ggsn)

data <- read.table("../../Data/iberian_bees.csv.gz", header = T, quote = "\"", sep = ",")

@@

@@

### 2022-07-01 Manuscript/manuscript.Rmd
manuscript update

Bees (Hymenoptera, Apoidea, Anthophila) are a diverse group of insects with more than 20,000 species described worldwide [@ascher2020]. The Iberian Peninsula, with its bee-loved Mediterranean climate, is one of the hotspots of bee biodiversity with more than 1,000 species recorded to date and counting [@ascher2020; @ortiz2020]. Unfortunately, despite the high biodiversity of Iberian bees, we know very little about the species themselves and their distributions. This is paradoxical, as in recent years bees have been highlighted as a keystone group in the ecosystem due to a key function they provide, the pollination of thousands of plants [@ollerton2011], including most crop species [@klein2007]. In addition, various rapid human-induced environmental changes are leading to the decline of some bee populations [@goulson2015]. However, the response of bee species to global change is heterogeneous both in space and also taxonomically. That is, while in some areas species are declining drastically [@burkle2013], other areas are well conserved [@herrera2019] and while some species are in the brink of extinction [@cameron2011] some species are even thriving [@russo2016]. A better description of bee distributions, habitat preferences, phenologies, or historical trends in the Iberian Peninsula would require information on when and where different species occur.

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',out.width="115%"}
library(sf)
library(raster)
library(rasterVis)
library(tidyverse)
library(data.table)
data <- read.table("../Data/iberian_bees.csv.gz", header = T, quote = "\"", sep = ",")

@@

@@

>>>>

Bees (Hymenoptera, Apoidea, Anthophila) are a diverse group of insects with more than 20,000 species described worldwide  (Ascher and Pickering, 2020). The Iberian Peninsula, with its bee-loved Mediterranean climate, is one of the hotspots of bee biodiversity with more than 1,000 species recorded to date and counting (Ascher and Pickering, 2020; Ortiz-Sánchez, 2020). Unfortunately, despite the high biodiversity of Iberian bees, we know very little about the species themselves and their distributions. This is paradoxical, as in recent years bees have been highlighted as a keystone group in the ecosystem due to a key function they provide, the pollination of thousands of plants [@ollerton2011], including most crop species [@klein2007]. In addition, various rapid human-induced environmental changes are leading to the decline of some bee populations [@goulson2015]. However, the response of bee species to global change is heterogeneous both in space and also taxonomically. That is, while in some areas species are declining drastically [@burkle2013], other areas are well conserved [@herrera2019] and while some species are in the brink of extinction [@cameron2011] some species are even thriving [@russo2016]. A better description of bee distributions, habitat preferences, phenologies, or historical trends in the Iberian Peninsula would require information on when and where different species occur.

@@

# Figura 1

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',out.width="115%"}
library(sf)
library(raster)
library(rasterVis)
library(tidyverse)
library(data.table)
library(ggplot2)
library(viridis)
library(ggspatial)
library(ggsn)

@@

# Figura 2

@@

# Figura 3
